---
title: "Combined Normalized Sales between 2018-19 for D-light bulbs"
author: "Daniel Obare"
date: "3/14/2020"
output: html_document
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# **Objective** 

## create a model that **predicts** sales for the D-light bulbs

Provided the data for sales of the D-light with various variables in play we model an **lm**
model to predict the sales for products


# package loading and data wrangling

Incudes checking for **NA's** and changing **data types** as well as cleaning the names

```{r}
require(magrittr) #piping operator 
require(dplyr) #data cleaning
require(tidyr) #wranglng
require(stringr) #strings
require(ggplot2) #plots
theme_set(theme_minimal())
load(file = "normalized.Rdata")
```


```{r}
normalize <- function(x) {
  return((x-min(x))/(max(x)-min(x)))
}

attach(comined_normalized)

colSums(is.na(comined_normalized))

normalized.sales <- comined_normalized %>%select(-Month) %>%  mutate(Quarter = factor(Quarter)) %>% 
  mutate(`Group Name` = factor(`Group Name`)) %>% mutate(`Product Name` = factor(`Product Name`)) %>% 
  mutate(`Group Name` = factor(`Group Name`)) %>% mutate(`Product Name` = factor(`Product Name`)) %>% 
  mutate(Class = factor(Class)) %>% 
  mutate(Revenue = normalize(Revenue)) %>%  mutate(`Unlock Price` = normalize(`Unlock Price`)) %>%
  mutate(X7 = normalize(X7)) %>% mutate(`income above 30,000` = as.numeric(`income above 30,000`)) %>% 
  mutate(`Upfront Price` = normalize(`Upfront Price`)) %>% 
  mutate(`Number of sales` = normalize(`Number of sales`)) %>% 
  janitor::clean_names()

```

# Model Fitting

## Normalizing and Sampling data

```{r}

attach(normalized.sales)
set.seed(4321)
index <- sample(1:nrow(normalized.sales), 0.8*nrow(normalized.sales))
train <- normalized.sales[index, ]
test <- normalized.sales[-index, ]

train <- train[which(complete.cases(train)), ]

```

## model only numerical data with Quarters 


```{r}

fit <- lm(number_of_sales~quarter+x7+revenue+kerosene_no+kerosene_yes+
            bank_account_no+bank_account_yes+does_the_client_own_their_home+does_the_client_rent+
            own_a_motorcycle_or_other_motorized_vehicle_no+own_a_motorcycle_or_other_motorized_vehicle_yes+
            gender_female+gender_male+income_above_30_000+income_between_10_000_and_30_000+
            income_less_than_10_000+stima_no+stima_yes+class, data = train)
AIC(fit)
BIC(fit)
s <- summary(fit)
pred <- predict(fit, test)
act_pred <- data.frame(cbind(actuals = test$number_of_sales, predicted = pred))
min_max_acc <- mean(apply(act_pred, 1, min)/apply(act_pred, 1, max))
```

## Model of the whole data

```{r}
fit1 <- lm(number_of_sales~., data = train)
AIC(fit1)
BIC(fit1)
s1 <- summary(fit1)
pred.1 <- predict(fit1, test)
act_pred.1 <- data.frame(cbind(actuals = test$number_of_sales, predicted = pred.1))
min_max_acc.1 <- mean(apply(act_pred.1, 1, min)/apply(act_pred.1, 1, max))
```

## Second Revised Model

```{r}

fit2 <- lm(number_of_sales~quarter+group_name+x7+revenue+kerosene_no+does_the_client_rent+does_the_client_own_their_home+
             gender_male+income_less_than_10_000+income_between_10_000_and_30_000+income_above_30_000+stima_yes, 
           data = train)
AIC(fit2)
BIC(fit2)
s2 <- summary(fit2)
pred.2 <- predict(fit2, test)
act_pred.2 <- data.frame(cbind(actuals = test$number_of_sales, predicted = pred.2))
min_max_acc.2 <- mean(apply(act_pred.2, 1, min)/apply(act_pred.2, 1, max))
```

## Third Revised Model

Check model **diagnostics** to prove it's sufficiency and accuracy

```{r}

fit3 <- lm(number_of_sales~quarter+group_name+x7+revenue+kerosene_no+does_the_client_own_their_home+
             gender_male+income_less_than_10_000+income_between_10_000_and_30_000+income_above_30_000+stima_yes, 
           data = train)
AIC(fit3)
BIC(fit3)
s3 <- summary(fit3)
pred.3 <- predict(fit3, test)
act_pred.3 <- data.frame(cbind(actuals = test$number_of_sales, predicted = pred.3))
min_max_acc.3 <- mean(apply(act_pred.3, 1, min)/apply(act_pred.3, 1, max))

```



## Revised model that provides sufficient evidence for predicton of Sales for Products 

```{r}
fit4 <- lm(number_of_sales~quarter+x7+revenue+kerosene_no+does_the_client_own_their_home+
             gender_male+income_less_than_10_000+income_between_10_000_and_30_000+income_above_30_000+stima_yes, 
           data = train)
AIC(fit4)
BIC(fit4)
s4 <- summary(fit4)
s4
pred.4 <- predict(fit4, test)
act_pred.4 <- data.frame(cbind(actuals = test$number_of_sales, predicted = pred.4))
min_max_acc.4 <- mean(apply(act_pred.4, 1, min)/apply(act_pred.4, 1, max))
min_max_acc.4

```

From **fit4** we find sufficient evidence to use this model for prediction:
     The **R^2** is close to 1 \n
     **Minimum-maximum accuracy** is approximately **72.7%** which is a bit higher compared to the other models \n
     The **Arkaike Information Criterion (AIC)** and **Bayesian Information Criterion (BIC)** are very low as well

```{r}
par(mfrow = c(2,2))
plot(fit4)

```

From the **residuals plot** points fun out which is a likely indicator that there is a relationship between **predictor** and **response** variable
There is **normality** aspect evidenced by the normal-QQ plot
All variables are **significant** evidenced by the significant stars

# Graphical Analysis and Tables of comparison

## Graphical analysis of the Total Combined Sales

```{r}
combo.sales <- combined_total %>% mutate(Quarter = factor(Quarter)) %>% 
  mutate(`Product Name` = factor(`Product Name`)) %>%
  mutate(`Group Name` = factor(`Group Name`)) %>%
  mutate(`Total sales` = normalize(`Total sales`)) %>% 
  mutate(`Unlock Price` = normalize(`Unlock Price`)) %>% 
  mutate(`Total Revenue in Ksh` = normalize(`Total Revenue in Ksh`)) %>% 
  janitor::clean_names()
  

table(group_name, product_name) %>% kableExtra::kable() %>% kableExtra::kable_styling()


combo.sales %>% ggplot(aes(total_revenue_in_ksh, total_sales, col = product_name, shapes = product_name))+
  geom_point()+
  facet_wrap(~group_name)+
  ggtitle("Total Sales as per Rvenue")

```

## Graphcal analysis of the Total 2018-19 Normalized data

```{r}
table(product_name, class) %>% kableExtra::kable() %>% kableExtra::kable_styling()


normalized.sales %>% ggplot(aes(quarter, number_of_sales))+
  geom_bar(stat = "identity", fill = "blue")+
  facet_wrap(~product_name)+ ggtitle("Number of Sales per Quarter in every Product")+
  xlab("2018-19 Quarters")+ ylab("Number of sales")+coord_flip()



normalized.sales %>% ggplot(aes(number_of_sales, color = class))+
  geom_histogram(fill = "white", position = "dodge")+
  facet_wrap(~product_name)+
  xlab("Number of Sales")+
  ggtitle("Sales as per Classes of Product")+
  theme(legend.position = "top")


normalized.sales %>% ggplot(aes(income_less_than_10_000))+
  geom_histogram(fill = "white", position = "dodge", color = "blue")+
  facet_wrap(~product_name)+
  xlab("Income less than 10,000 KShs")+
  ggtitle("Income less 10,000 KShs as per Product")


normalized.sales %>% ggplot(aes(income_between_10_000_and_30_000))+
  geom_histogram(fill = "white", position = "dodge", color = "purple")+
  facet_wrap(~product_name)+
  xlab("Income between 10,000 - 30,000 KShs")+
  ggtitle("Income between 10,000 - 30,000 KShs as per Product")+
  theme(legend.position = "top")



normalized.sales %>% ggplot(aes(income_above_30_000))+
  geom_histogram(fill = "white", position = "dodge", color = "green")+
  facet_wrap(~product_name)+
  xlab("Income above 30,000 KShs")+
  ggtitle("Income above 30,000 KShs as per Product")+
  theme(legend.position = "top")


normalized.sales %>% ggplot(aes(stima_no))+
  geom_histogram(fill = "white", position = "dodge", color = "blue")+
  facet_wrap(~product_name)+
  xlab("Clients with no Stima")+
  ggtitle("Clients with no Stima as per Product")+
  theme(legend.position = "top")


normalized.sales %>% ggplot(aes(stima_yes))+
  geom_histogram(fill = "white", position = "dodge", color = "green")+
  facet_wrap(~product_name)+
  xlab("Clients with Stima")+
  ggtitle("Clients with Stima as per Product")+
  theme(legend.position = "top")

normalized.sales %>% ggplot(aes(gender_female, number_of_sales))+
  geom_point(color = "blue")+
  facet_wrap(~product_name)+
  xlab("Female Clients")+
  ylab("Number of Sales")+
  ggtitle("Female Sales")

normalized.sales %>% ggplot(aes(gender_male, number_of_sales))+
  geom_point(color = "purple")+
  facet_wrap(~product_name)+
  xlab("Male Clients")+
  ylab("Number of Sales")+
  ggtitle("Male Sales")
```









